---
title: "p8105_fp_ds100"
author: "Yue Gu, Jianghui Lin, Junyuan Zheng, Jianyou Liu, Zhiqian Fang"
date: "11/5/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
library(tidyverse)
```

## Proposal 

#### The group members (names and UNIs) 
```{r, echo = FALSE}
names = c("Yue Gu", "Jianghui Lin", "Junyuan Zheng", "Jianyou Liu", "Zhiqian Fang")
uni = c("yg2625", "jl5172", "jz3036", "jl5296", "zf2212")
rbind(names, uni) %>% 
  knitr::kable()
```

#### The tentative project title

Title: __1999 - 2016 U.S. Injury Mortality Analysis__

#### The motivation for this project

Motivation: 
Suicide is a leading cause of death in the US. Suicide rates increased in nearly every state from 1999 through 2016. Mental health conditions are often seen as the cause of suicide, but suicide is rarely caused by any single factor. In fact, many people who die by suicide are not known to have a diagnosed mental health condition at the time of death.

reference: “Suicide Rising across the US.” Centers for Disease Control and Prevention, Centers for Disease Control and Prevention, 11 June 2018, www.cdc.gov/vitalsigns/suicide/index.html.


#### The intended final products

We will first present the trend of injury intent from 1999 to 2016 over all ages and races. Then, we will focus on suicide mortality. The final product of this project would be presented as a website containing discription of our motivation, data source, statistical methods, and conclusion. Barplot plots would be shown to demonstrate suicide potential trend according to sex, age, race, and injury mechanism across years.

#### The anticipated data sources
https://data.cdc.gov/NCHS/NCHS-Injury-Mortality-United-States/nt65-c7a7

#### The planned analyses / visualizations / coding challenges
We aim to analyze the suicide rate changes from 1999-2016 across different ages/races.
#### The planned timeline




There should be one proposal per group, written collaboratively using .Rmd (rendering to a GH document) in a dedicated GitHub repo. Conceptually, this is intended to review a project that is 10% complete.






