---
title: "p8105_fp_ds100"
author: "Yue Gu, Jianghui Lin, Junyuan Zheng, Jianyou Liu, Zhiqian Fang"
date: "11/5/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
library(tidyverse)
```

## Proposal 

#### The group members (names and UNIs) 
```{r, echo = FALSE}
names = c("Yue Gu", "Jianghui Lin", "Junyuan Zheng", "Jianyou Liu", "Zhiqian Fang")
uni = c("yg2625", "jl5172", "jz3036", "jl5296", "zf2212")
rbind(names, uni) %>% 
  knitr::kable()
```

#### The tentative project title

Title: __1999 - 2016 U.S. Injury Mortality Analysis__

#### The motivation for this project

Motivation: 

#### The intended final products
#### The anticipated data sources
#### The planned analyses / visualizations / coding challenges
#### The planned timeline

There should be one proposal per group, written collaboratively using .Rmd (rendering to a GH document) in a dedicated GitHub repo. Conceptually, this is intended to review a project that is 10% complete.






