---
title: "p8105_fp_ds100_yg2625"
author: "Yue Gu"
date: "11/30/2018"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
# Regression Model Analysis

#### Load BRFSS & Injury Data
```{r read_data, message = FALSE}
brfss_data = read_csv("./data/brfss_data.csv") %>% 
  janitor::clean_names()

injury_data = read_csv("./data/NCHS_-_Injury_Mortality__United_States.csv") %>% 
  janitor::clean_names()

```

## Suicide rate model
#### Convert variables into 'factor' to fit a linear model
```{r convert_var}
new_injury = 
  injury_data %>% 
  filter(injury_mechanism == "All Mechanisms",
         injury_intent == "Suicide",
         year  %in%  c(2011, 2012, 2013, 2014, 2015, 2016),
         age_group_years != "All Ages",
         race != "All races",
         sex != "Both sexes") %>%
  mutate(suicide_dr = (deaths / population) * 100000,
         age_group_years = str_replace(age_group_years, "< 15", "<25"),
         age_group_years = str_replace(age_group_years,"15–24", "<25"),
         age_group_years = str_replace(age_group_years,"65–74", "65+"),
         age_group_years = str_replace(age_group_years,"75+", "65")) %>% 
  select(year, sex, age_group_years, race, suicide_dr) %>% 
  mutate(sex = as.factor(ifelse(sex == "Male", 0, 1)),
         race = as.factor(race),
         age_group_years = as.factor(age_group_years),
         year = as.factor(year))
  
```
#### Construct a regression model using 'suicide death rate' as the outcome of interest with 'sex', 'age group' and 'race' being its predictors
```{r lm}
fit_suicide = lm(suicide_dr ~ sex + race + age_group_years + year, data = new_injury)
summary(fit_suicide)

pairwise.t.test(new_injury$suicide_dr, new_injury$age_group_years, p.adj = 'bonferroni')

fit_suicide %>% aov(.) %>%  TukeyHSD()
```


## Depression model

#### Age model
```{r}

new_brfss_age =
  brfss_data %>% 
  filter(year %in% c(2011, 2012, 2013, 2014, 2015, 2016),
         response == "Yes",
         age_group != "",
         data_value != "") %>%
  mutate(age_group = str_replace(age_group, "18-24", "<25"),
         age_group = str_replace(age_group, "25-34", "25-44"),
         age_group = str_replace(age_group,"35-44", "25-44"),
         age_group = str_replace(age_group,"45-54", "45-64"),
         age_group = str_replace(age_group,"55-64", "45-64")) %>% 
  select(data_value, age_group) %>% 
  mutate(age_group = as.factor(age_group))

fit_age = lm(data_value ~ age_group, data = new_brfss_age)
summary(fit_age)

fit_age %>% aov(.) %>% TukeyHSD()
```

#### Gender model

```{r}
new_brfss_gender =
  brfss_data %>% 
  filter(year %in% c(2011, 2012, 2013, 2014, 2015, 2016),
         response == "Yes",
         gender != "",
         data_value != "") %>%
  mutate(total_sp = sample_size/(data_value/100)) %>% 
  select(data_value, gender) %>% 
  mutate(gender = as.factor(ifelse(gender == "Male", 0, 1)))


fit_gender = lm(data_value ~ gender, data = new_brfss_gender)
summary(fit_gender)

fit_gender %>% aov(.) %>% TukeyHSD()
```


#### Race model

```{r}
new_brfss_race =
  brfss_data %>% 
  filter(year %in% c(2011, 2012, 2013, 2014, 2015, 2016),
         response == "Yes",
         race_ethnicity == "Black, non-Hispanic" | race_ethnicity == "White, non-Hispanic" | race_ethnicity == "Hispanic",
          data_value != "") %>%
  select(data_value, race_ethnicity) %>% 
  mutate(race_ethnicity = as.factor(race_ethnicity),
         race_ethnicity = fct_relevel(race_ethnicity, race_ethnicity = c("Hispanic", "Black, non-Hispanic", "White, non-Hispanic")))


fit_race = lm(data_value ~ race_ethnicity, data = new_brfss_race)
summary(fit_race)

fit_race %>% aov(.) %>% TukeyHSD()

```




## Analysis Report
To maintain comparability between injury mechanism and depression(brfss) data, we only generate statistic results for:  

* race:
    + Hispanic*(baseline)*
    + Non-Hispanic Black
    + Non-Hispanic White
    
* age groups(in years):
    + < 25*(baseline)*
    + 25-44
    + 45-64
    + 65+
* sex/gender:
    + = 0 if male*(baseline)*
    + = 1 if female
* year: 
    + 2011*(baseline)*
    + 2012
    + 2013
    + 2014
    + 2015
    + 2016

### Suicide Model
Using injury data, we first calculate suicide death rate by
      
$$  \frac{Deaths}{Population} * 100000
$$
    which represents the number of deaths caused by suicide per 100,000 units population. Then, we transform sex, race, age, year into factor variables for future regression model building. 

Then, we construct regression for suicide death rate as response, sex, race, age(in years), year as predictor. By observing the coefficients estimates and p-values, there is several interesting finds for suicide death rates:    
1. sex has **significant** p-value < 0.0001: there is significant difference of suicide rate between male and female group and female tends to have lower suicide death rate.   
2. race of Non-Hispanic white has **significant** p-value: there is significant difference of suicide rate between Non-hispanic Whites and Hispanic and Non-hispanic Black have non-significant difference with Hispanic. Whites tend to have higher suicide death rate compared to Hispanic and Blacks.   
3. all groups of age(in years) have **significant** p-value: there is significant difference between all age groups with baseline age group(age < 25).    
4. all years have **non-significant** p-value: there is not significant difference between all years with year 2011.   
5. the model produced an adjusted $R^2$ = 0.6405, which represents there are 64% of the variability of the suicide rate are explained by the fitted model and data after adjusted and it's a acceptable proportion for the model.

Then, we make pairwise comparison with Bonferroni and Tukey for race, age and year groups. Findings:      
1.  __Race__: Non-Hispanic White and Hispanic, Non-Hispanic White and Black have significant different suicide death rate. White have the highest suicide death rate among 3 race groups. And Blacks have smaller suicide death rate compared to Hispanic and Whites.  
2.  __Age__: 25-44 and <25, 45-64 and <25, 65+ and <25 have significant different suicide death rate. Age groups that are >25 all have higher suicide death rate compared to age <25. And age 65+ have smaller suicide death rate compared to 25-44, 45-64 groups.    
3. __Year__: All pairwise comparison for years don't generate significant result, meaning there is no significant different suicide death rate in different years.

 
### Depression Model
Because of the data structure of brfss data, the age, gender, race are independent characteristics of the participants to the study, we have to build 3 seperate model for independent analysis. And the data_value(in %) represents the proportion of people have depression.

**Age Model**:    
1.  There is significant difference in the depression propotion between age group <25 and each other age group including 25-44, 45-64, 65+.   
2.  All pairwise comparison showed a significant p-value between each age groups while 25-44, 45-64 ages showed an increased depression proportion and 65+ showed a decreased depression proportion.
3.  The adjusted $R^2$ = 0.2617 indicates that 
26.17% of variability of the depression proportion is explained by the model only includes age groups as predictor.

**Gender Model**:   
1.  There is significant difference in the depression propotion between male and female. And female indicates a higher depression proportion than male.  
2.  The adjusted $R^2$ = 0.6146 indicates that 
61.46% of variability of the depression proportion is explained by the model only includes gender as predictor.

**Race Model**:   
1.  There is significant difference in the depression propotion between Hispantic and Non-Hispanic Black, Hispantic and Non-Hispanic White.   
2.  All pairwise comparison showed a significant p-value between each race groups while Whites have higher depression proportion than Hispanics and Blacks and Blacks have lower depression proportion than Hispanics.
3.  The adjusted $R^2$ = 0.1081 indicates that 
10.81% of variability of the depression proportion is explained by the model only includes race groups as predictor.


### Additional Analysis for Location:

**Location Model**:   

#### Location
```{r}
new_brfss_location =
  brfss_data %>% 
  filter(year %in% c(2011, 2012, 2013, 2014, 2015, 2016),
         data_value != "") %>%
  select(locationabbr, data_value) %>% 
  mutate(locationabbr = as.factor(locationabbr))

fit_location = lm(data_value ~ locationabbr, data = new_brfss_location)
summary(fit_location)

```
From model output, we observe that almost all p-value except for VI are non-significant. However, the model generates a negative adjusted $R^2$, which indicates that the explanation towards depression proportion is low or neglibible and it shows the insignificance of explanatory variables for location as predictors, thus, we can't get analytical result for location in the data.
